---
title: "p8105_hw1_sc5826"
author: "Shivalika Chavan"
date: "2025-09-11"
output: github_document
---
**Problem 1**
```{r, echo = TRUE, message = FALSE}
library(tidyverse)
library(moderndive)
data("early_january_weather")
```

The `early_january_weather` dataset has `r ncol(early_january_weather)` variables (`r paste(colnames(early_january_weather), sep = ",")`). It contains weather data originating from `r unique(early_january_weather$origin)`, between days `r min(early_january_weather$day)` and `r max(early_january_weather$day)` of the `r unique(early_january_weather$month)`st month of `r unique(early_january_weather$year)` (January 1-15, 2013). Data  was recorded between `r min(early_january_weather$time_hour)` and `r max(early_january_weather$time_hour)`. There are `r nrow(early_january_weather)` total observations during this period. 

With temperature ranging between `r min(early_january_weather$temp)` and `r max(early_january_weather$temp)`, we can assume these were in Fahrenheight. The mean temperature observed in this period was `r round(mean(early_january_weather$temp), 2)`F. 

```{r echo = FALSE}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  ylab("Temperature (F)") 
ggsave("temp_humidity_scatter_plot.pdf", height = 4, width = 6)
```

There is a cyclical pattern observed in temperature, where the weather appears to peak mid-day and then falls during the night. The peaks appear to trend upward (temperature is getting warmer), until January 14th, where it steeply drops further down. Right before the steep drop in temperature, there also is a sharp increase in the humidity to `r max(early_january_weather$humid)`. This is higher than the average humidity during this period of observation (`r round(mean(early_january_weather$humid), 0)`). 

**Problem 2**
```{r collapse = TRUE}
set.seed(123)
df_problem_2 = tibble(
  vec_numeric = rnorm(10),
  vec_logical = vec_numeric>0,
  vec_char = vector(mode="character", length=10),
  vec_factor = factor(rep(NA, 10), levels = c("A", "B", "C"))
)

mean(pull(df_problem_2, vec_numeric))
mean(pull(df_problem_2, vec_logical))
mean(pull(df_problem_2, vec_char))
mean(pull(df_problem_2, vec_factor))
```
Taking the mean of the numeric and logical vector works. The logical vector works because it can be implicitly converted to 0s and 1s, which are numbers. Neither the character or factor vectors can be converted into a numeric or logical vector. R then throws an error and marks the result as `NA`. 

```{r echo = TRUE, results = "hide"}
as.numeric(pull(df_problem_2, vec_logical))
as.numeric(pull(df_problem_2, vec_char))
as.numeric(pull(df_problem_2, vec_factor))
```
The logical vector is coerced into a numeric vector of 0s and 1s. When the character and factor vectors are coerced into a numeric vector, each element becomes NA because characters and factors do not have an inherent numeric value. This is unlike logical vectors where `TRUE = 1` and `FALSE = 0`. If a vector cannot be numeric (or logical), then the mean cannot be calculated. 


