---
title: "p8105_hw1_sc5826"
author: "Shivalika Chavan"
date: "2025-09-11"
output: github_document
---
**Problem 1**
```{r, echo = TRUE, message = FALSE}
library(tidyverse)
library(moderndive)
data("early_january_weather")
```

The `early_january_weather` dataset has `r ncol(early_january_weather)` variables (`r paste(colnames(early_january_weather), sep = ",")`). It contains weather data originating from `r unique(early_january_weather$origin)`, between days `r min(early_january_weather$day)` and `r max(early_january_weather$day)` of the `r unique(early_january_weather$month)`st month of `r unique(early_january_weather$year)` (January 1-15, 2013). Data  was recorded between `r min(early_january_weather$time_hour)` and `r max(early_january_weather$time_hour)`. There are `r nrow(early_january_weather)` total observations during this period. 

With temperature ranging between `r min(early_january_weather$temp)` and `r max(early_january_weather$temp)`, we can assume these were in Fahrenheight. The mean temperature observed in this period was `r round(mean(early_january_weather$temp), 2)`F. 

```{r echo = FALSE}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  ylab("Temperature (F)") 
ggsave("temp_humidity_scatter_plot.pdf", height = 4, width = 6)
```
There is a cyclical pattern observed in temperature, where the weather appears to peak mid-day and then falls during the night. The peaks appear to trend upward (temperature is getting warmer), until January 14th, where it steeply drops further down. Right before the steep drop in temperature, there also is a sharp increase in the humidity to `r max(early_january_weather$humid)`. This is higher than the average humidity during this period of observation (`r round(mean(early_january_weather$humid), 0)`). 




